Product graphics

### 2009-11-28 product-graphics.tex
More working

  \item {\bf spine}: width is proportional to value, height equally divides space. Splines are space filling, but make it more difficult to compare absolute counts.  Splines can be arranged horizontally (``hspline'') or vertically (``vspline'').

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.33\linewidth]{part-hbar-vr}%
    \includegraphics[width=0.33\linewidth]{part-hspline-vr}
    % \includegraphics[width=0.25\linewidth]{part-hspline-vr}

  \caption{1d partitions showing proportions 0.5, 0.2, 0.2 and 0.1. From left to right: vertical bars, horizontal bars, vertical splines and horizontal splines. }
  \label{fig:part-1d}
\end{figure}

@@

One special 2d case is the equal bin size plot - which all 1d x 1d primitives and the fluct collapse to when counts are equal.

@@

The sequence of partitioning choices defines the different named area plots:

\begin{itemize}
  \item {\bf Stacked} barchart: 1 bar + $n-1$ spines in opposite direction.
  \item {\bf Nested} barchart: $n$ bars.
  \item {\bf Mosaic} plot: spines in alternating directions.  The 1d case has a special name, the spineplot. \citep{hartigan:1984,hartigan:1981,friendly:1994,hofmann:2003}
  \item {\bf Double-decker} plot: $n-1$ spines + 1 spine in opposite direction.

  \item {\bf Fluctuation} diagram: fluct, then spline. 
  \item {\bf Equal bin size} plot: a fluctuation diagram, when the 2 variables have uniform distribution.
\end{itemize}

>>>>

  \item {\bf spine}: width is proportional to value, height equally divides space. Splines are space filling, but make it more difficult to compare absolute counts.  Splines can be arranged horizontally (``hspline''),  vertically (``vspline''), or can automatically pick their orientation (``spline'').

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.33\linewidth]{part-hbar}%
    \includegraphics[width=0.33\linewidth]{part-hspline}
    % \includegraphics[width=0.25\linewidth]{part-treemap}

  \caption{1d partitions showing proportions 0.5, 0.2, 0.2 and 0.1. From left to right: horizontal bars, horizontal splines, and a treemap.}
  \label{fig:part-1d}
\end{figure}

@@

One special 2d case is the equal bin size plot - which all 1d x 1d primitives and the fluct collapse to when proportions are equal, as when conditioned.

Figure~\ref{fig:fluct} shows these two behaviours, looking at the relationship between happiness, health and financial status. On the left, a display of the raw proportion shows that most people are in good health and average financial standing, but it is difficult to see how happiness varies because the size of the squares is variable. On the right, conditioning on financial status and health makes all squares the same size and makes it easier to see the conditional distribution of happiness - people tend to be happier the healthier and the richer they are

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\linewidth]{part-fluct}%
    \includegraphics[width=0.5\linewidth]{part-fluct-cond}
  \caption{caption}
  \label{fig:fluct}
\end{figure}

@@

The sequence of partitioning choices defines different named area plots:

\begin{itemize}
  \item {\bf Barchart} (1d). 1 hbar.
  \item {\bf Column chart} (1d). 1 vbar.
  \item {\bf Spineplot} (1d). 1 vspline.

  \item {\bf Stacked} barchart (2d). 1 hbar and 1 vspine.
  \item {\bf Nested} barchart (2d).  2 hbars. \citep{peltier:2009}
  \item {\bf Fluctuation} diagram (2d): fluct.

  \item {\bf Equal bin size} plot (3d): fluct and vspline, provided the final two variables are uniform.

  \item {\bf Mosaic} plot (nd).  Alternating hspines and vspines.  \citep{hartigan:1984,hartigan:1981,friendly:1994,hofmann:2003}
  \item {\bf Double-decker} plot (nd).  $n-1$ hspines and 1 vspine. \citep{hofmann:2001}
  \item {\bf Treemap} (nd): n spines. \citep{shneiderman:1992}
  \item {\bf Squarified treemap} (nd): n squares. \citep{bruls:1999}

% http://peltiertech.com/Utility/ClusterStackUtility.html

\begin{itemize}
